---
title: "Homework 6"
author: "mk4995"
date: "2024-11-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)

set.seed(1)
```

## Problem 1

This problem uses the Central Park data that we used earlier in the semester from `rnoaa`.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


Start by conducting a simple linear regression (tmax ~ tmin) as a basic test.

```{r}
fit = lm(tmax ~ tmin, data = weather_df)

fit |> broom::tidy()
fit |> broom::glance()
```

Do some bootstrapping.

```{r}
boot_straps = 
  weather_df |> 
  modelr::bootstrap(5000) |> 
  mutate(
    strap = map(strap, as_tibble), 
    models = map(strap, \(df) lm(tmax ~ tmin, data = df)),
    tidy_results = map(models, broom::tidy),
    glance_results = map(models, broom::glance)
  ) |> 
  select(.id, tidy_results, glance_results)
```

Gather results.

```{r}
bootstrap_r2 =
  boot_straps |> 
  unnest(glance_results) |> 
  group_by(.id) |> 
  select(.id, r.squared)

bootstrap_log = 
  boot_straps |> 
  unnest(tidy_results) |> 
  select(.id, term, estimate) |> 
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  ) |> 
  rename(
    b0 = `(Intercept)`,
    b1 = tmin
  ) |> 
  mutate(log_beta_product = log(b0*b1)) |> 
  select(.id, log_beta_product)

bootstrap_results = 
  full_join(bootstrap_r2, bootstrap_log, by = ".id") |> 
  ungroup()

bootstrap_results |> 
  head(20) |> 
  knitr::kable()
```

The table above shows the first 20 rows of data.

```{r}
bootstrap_results |> 
  ggplot(aes(x = r.squared)) + geom_histogram() + theme_minimal()
```

The distribution of the `R-squared` values from bootstrapping 5000 samples is relatively normal with the center of the distribution around 0.91. Almost all values are contained between 0.88 and 0.94.

```{r}
bootstrap_results |> 
  ggplot(aes(x = log_beta_product)) + geom_histogram() + theme_minimal()
```

The distribution of the `log(B1 * B2)` values from bootstrapping 5000 samples is also relatively normal with the center of the distribution around 2.02. Almost all values are contained between 1.94 and 2.08.

Create confidence intervals.

```{r}
bootstrap_results |> 
  dplyr::summarize(
    r_squared_mean = mean(r.squared),
    r_squared_ci_ll = quantile(r.squared, 0.025),
    r_squared_ci_ul = quantile(r.squared, 0.975),
    log_beta_product_mean = mean(log_beta_product),
    log_product_ci_ll = quantile(log_beta_product, 0.025),
    log_product_ci_ul = quantile(log_beta_product, 0.975)
  ) |> 
  knitr::kable()
```


